---
title: "California Bargaining Unit 6 Compensation Analysis"
author: "Don Boyd"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_notebook:
    df_print: paged
    fig_height: 6
    fig_width: 8
    toc: yes
    toc_depth: 3
    toc_float: true
    number_sections: yes
editor_options:
  chunk_output_type: console
---


```{r links}
# https://download.bls.gov/pub/time.series/cm/
# NCS: https://download.bls.gov/pub/time.series/nb/
# https://www.pewtrusts.org/~/media/Assets/2014/08/StateEmployeeHealthCareReportSeptemberUpdate.pdf

```


```{r runall, eval=FALSE, echo=FALSE} 
# When we want a final report, run the following code selectively "by hand" (interactively)
# -- NEVER using Knit with eval=TRUE
# note <br> breaks line for html output but \n should break for pdf; also could try |

rmdfn <- ".CA_report.rmd" # this file
outfn <- paste0("report_", format(Sys.time(), "%Y-%m-%d"), ".html")
rmarkdown::render(rmdfn, output_format="html_document", output_file=outfn)

# library("RCurl")
# ftpUpload(outfn, "ftp://kffrig:boyd4812@files.000webhost.com/public_html/KFF_RIG_MedicaidCuts_new.html")

# Note may b safest to fully exit RStudio and restart it before running whole thing. Otherwise knitr may get confused
# and include repetitive information in the output html file.

```


```{r setup, include=FALSE, eval=TRUE}
knitr::opts_chunk$set(echo = FALSE, include=FALSE, eval=TRUE, message = FALSE, warning = FALSE)
```


```{r libraries}

library(tidyverse)
options(tibble.print_max = 80, tibble.print_min = 80) # if more than 60 rows, print 60 - enough for states
library(purrr)
library(haven)
library(RSQLite)

library(Hmisc)

library(kableExtra)
library(gt)
library(knitr)
library(arrow)

library(maps)
# https://cran.r-project.org/web/packages/usmap/vignettes/mapping.html
library(usmap)
library(gridExtra)
library(RcppRoll)
library(ggbreak)
library(patchwork)
library(RColorBrewer)

library(fixest)

library(btools)
library(bdata)
```


```{r locations}
# acsdir <- r"(E:\data\acs\)"
# pumsdir <- paste0(acsdir, "pums/")
# csvdir <- paste0(pumsdir, "csv/")

acsdir <- r"(E:\data\acs\pums\sqlite\)"
acsfn <- "acs2019_5year"
acsdb <- paste0(acsdir, acsfn, ".sql3")

```


```{r constants}
# 33-3012	Correctional Officers and Jailers  OES code
# 3802 Correctional officers and jailers 33-3012   # occ code 2018
# NY, MA
# TX, FL
# AZ, NV, OR, WA

compstates <- c("CA", "AZ", "NV", "OR", "WA", "TX", "FL", "NY", "MA")


# sts <- c("CA", "NY", "FL", "NV", "TX")
# storder <- c("other", sts)
# 
# sts <- c("CA", "MA", "RI", "CT", "NY", "FL", "NV", "TX") %>% sort
# storder <- c("other", sts)

```


```{r load_data}
load(file=here::here("data", "caloes.rdata"), verbose=TRUE)

```


```{r functions}
get_stname <- function(stabbr){
  sts <- c("US", "DC", state.abb)
  allnames <- c("United States", "District of Columbia", state.name)
  allnames[match(stabbr, sts)]
}

```

# What data for what purpose


# Comparisons to Corrections Officers in Other States

## OES data
```{r oes_notes}
# https://www.bls.gov/oes/oes_con.htm
# Two types of OEWS estimates can be downloaded directly from this website in Excel (XLS) format:
# 
# Occupational employment and wage estimates by geographic area (national, state, metropolitan, and nonmetropolitan area)
# National Occupational employment and wage estimates by industry (sector, 3-, 4-, and selected 5- and 6-digit North American Industry Classification System levels)
# Download Occupational Employment and Wage Estimates
# 
# In addition to the estimates that can be downloaded directly from this website, historical OEWS estimates are available upon request through the contact information above. Historical OEWS estimates consist of the 1988 to 1995 national occupational employment and wage data by industry, for most industries at the 2- and 3-digit Standard Industrial Classification (SIC) levels. Estimates for 1988 - 1995 contain only employment data. Between 1988 and 1995 covered industries were surveyed once in a three-year cycle. The OEWS program now surveys all covered industries each year.
# 
# Note: 1996 estimates are not available.
# 
# OEWS State Employment and Wage Estimates
# Occupational Employment and Wage Statistics estimates, by occupation and by industry, for individual states are available from the states' Labor Market Information (LMI) or Research, Analysis, and Statistics offices which are part of their State Employment Security Agencies (SESAs). Availability, format and medium of the data varies by state. To obtain OEWS estimates for a particular state, please contact the appropriate office on the State Contact List.
```



```{r}
# main oes data
# Hi, I'm looking for OES data for states, and possible metro areas, ideally in a single file. I have found the year-by-year state Excel files (https://www.bls.gov/oes/tables.htm) which are awkward to work with but not a bad last resort, and I have found the link for the new OEWS-titled data (https://download.bls.gov/pub/time.series/oe/) but that is only for the latest release. I was hoping you have the historical data in a single csv or flat ASCII file, for all years and states (and maybe metro areas). Would you have a link to a download site for these data? Many thanks in advance. Don Boyd

```



```{r bls_research}
#.. BLS state research estimates ----
# https://www.bls.gov/oes/current/oes_research_estimates.htm
# The OEWS program provides wage and employment estimates by state and industry
# beginning with the May 2012 reference period. These estimates are intended for
# research purposes, and users should be aware of the limitations of the data.
# Estimates prior to May 2012 are not available. Estimates are only available
# for states; industry-specific estimates for metropolitan areas are not
# available. Estimates are only available for detail and major occupation
# groups; industry-specific estimates for broad and minor occupation groups are
# not available.
oes <- readRDS(paste0(resdir, "oesres_sec99.rds"))
glimpse(oes)
count(oes, own_code, o_group)
count(oes, naics_title)

# get 1st-line supervisors and corrections officers
# 33-1011	First-Line Supervisors of Correctional Officers	High school diploma or equivalent
# 33-3012	Correctional Officers and Jailers	High school diploma or equivalent
corr <- oes %>%
  filter(occ %in% c("33-1011", "33-3012"),
         !str_detect(naics_title, "Federal")) %>%
  mutate(level=case_when(str_detect(naics_title, "State") ~ "state",
                         str_detect(naics_title, "Local") ~ "local",
                         TRUE ~ "ERROR"))
count(corr, level, own_code, o_group)
count(corr, naics_title)
count(corr, year, level) %>% pivot_wider(names_from = level, values_from = n)
corr %>% filter(stabbr=="CA", year==2012)
corr %>% filter(is.na(stabbr))
count(corr, stabbr)
summary(corr)  # annual has no data

corr2 <- corr %>%
  filter(stabbr %in% state.abb) %>%
  select(year, occ, occname, stabbr, area=area_title, level, emp=tot_emp, awage=a_mean, awage_mdn=a_median)

corr2 %>% filter(stabbr=="CA")


```


### Graphs and tables of OES research
```{r}

corr2 %>%
  filter(year==2020) %>%
  select(year, occ, occname, stabbr, level, awage) %>%
  pivot_wider(names_from = level, values_from = awage) %>%
  group_by(occ) %>%
  mutate(across(c(state, local), list(rank = ~ rank(-.x)))) %>%
  ungroup %>%
  arrange(occ, state_rank)

cols <- c('#a6cee3','#1f78b4','#b2df8a','#33a02c','#fb9a99','#e31a1c','#fdbf6f','#ff7f00','#cab2d6')
cols <- c('#a6cee3','#1f78b4','black','#33a02c','#fb9a99','#e31a1c','#fdbf6f','#ff7f00','#cab2d6')

p <- corr2 %>%
  filter(occ=="33-3012", stabbr %in% compstates, level=="state") %>%
  mutate(cadum=ifelse(stabbr=="CA", 1, 0) %>% as.factor) %>%
  arrange(desc(cadum), area) %>%
  ggplot(aes(year, awage / 1000, colour=area)) +
  geom_line(size=1.0) +
  geom_point() +
  scale_colour_manual(values=cols) +
  # scale_size_manual(values=c(1, rep(.5, 8))) +
  labs(y="Average annual wage ($thousands)", x=NULL,
       caption="Source: U.S. Bureau of Labor Statistics, Occupational Employment and Wage Statistics") +
  scale_y_continuous(breaks=seq(0, 200, 10), labels = scales::dollar, limits=c(0, NA)) +
  ggtitle("Average annual salaries of state correctional officers",
          subtitle="California, neighbors, and selected other states") +
  theme_bw() +
  theme(legend.title = element_blank(),
        plot.caption = element_text(hjust = 0))
p
ggsave(here::here("results", "oes_wages.png"), plot = p,
       width=10, height=8, scale=1)


```




## Intra-California
```{r}
load(file=here::here("data", "caloes.rdata"), verbose=TRUE)  # just California
glimpse(cal_oes)
count(cal_oes, atype)
count(cal_oes , aname)

```


## Education requirements
```{r}
# https://www.bls.gov/oes/additional.htm
# The typical education level required to enter an occupation is based on education and training categories from the BLS Employment Projections program. Because of changes to the occupational classification system that began with the May 2019 OEWS estimates and because the entry-level educational requirements assigned to some occupations have changed, the May 2020 OEWS data by typical-entry level educational requirement are not directly comparable to OEWS typical entry-level education data prior to 2019. The May 2020 estimates are based primarily on the system of education and training requirements used for the 2018-2028 employment projections.

# https://www.bls.gov/oes/2020/may/education_2020.xlsx
# 33-1011	First-Line Supervisors of Correctional Officers	High school diploma or equivalent
# 33-3012	Correctional Officers and Jailers	High school diploma or equivalent

# ONETIME
# ujobed <- "https://www.bls.gov/oes/2020/may/education_2020.xlsx"
# download.file(url=ujobed, destfile=here::here("data_raw", "education_2020.xlsx"), mode="wb")
# END ONETIME

jobed1 <- read_excel(here::here("data_raw", "education_2020.xlsx"), sheet = "educ_list", skip=3)
jobed <- jobed1 %>%
  setNames(c("occ", "occname", "entryed"))
count(jobed, entryed)
jobed %>%
  filter(str_sub(occ, 1, 2)=="33")

# get 2020 records in CA for HS jobs
hsjobs <- jobed %>%
  filter(entryed=="High school diploma or equivalent")

oeshs1 <- readRDS(paste0(resdir, "oesres_allsectors.rds")) %>%
  filter(stabbr=="CA", year==2020, occ %in% hsjobs$occ)

oeshs2 <- oeshs1 %>%
  filter(tot_emp >= 1000, i_group=="sector") %>%
  select(stabbr, year, occ, occname, naics, naics_title, i_group, tot_emp, h_mean, a_mean) %>%
  arrange(-a_mean)


oeshs3 <- oeshs1 %>%
  filter(tot_emp >= 10000, i_group=="sector", 
         !str_detect(occname, coll("supervisor", ignore_case = TRUE))) %>%
  select(stabbr, year, occ, occname, naics, naics_title, i_group, tot_emp, h_mean, a_mean) %>%
  arrange(-a_mean)

# collapse by occupation ----
oeshs_mean <- oeshs2 %>%
  filter(i_group=="sector", 
         !str_detect(occname, coll("supervisor", ignore_case = TRUE))) %>%
  mutate(totwage=tot_emp * a_mean) %>%
  group_by(stabbr, year, occ, occname) %>%
  summarise(tot_emp=sum(tot_emp), totwage=sum(totwage), .groups = "drop") %>%
  mutate(awage=totwage / tot_emp) %>%
  arrange(-awage)
summary(oeshs_mean)

# Most criminal investigators and detectives earn an associate degree in criminal justice or law enforcement, and some earn a bachelor's degree. Most graduate from a police training academy and work as police officers to gain law enforcement experience. A promotion is often required to become a criminal investigator.


# https://www.cdcr.ca.gov/por/minimum-qualifications-2/minimum-qualifications/
# Graduation from a U.S. high school, G.E.D. or equivalent from a U.S. institution, or a California High School Proficiency Examination (CHSPE) certificate is required, or possession of a college degree (Associate of Arts or higher) from an accredited college or university.

```



# Ideas

*   Top private sector occs with hs grad in CA vs corrections in CA
*   Top publics us vs ca, comp to bu6 differential


# Descriptive statistics from Cal OES
```{r}
count(cal_oes, wtype)
count(cal_oes, atype)


base <- 2009
dflong <- cal_oes %>%
  filter(soc=="333012", wtype=="annual", atype=="state") %>%
  pivot_longer(cols=wage:wp90) %>%
  group_by(name) %>%
  mutate(ivalue=value / value[year==base] * 100 - 100,
         cagr2009=(value / value[year==2009])^(1 / (year - 2009)) * 100 - 100,
         cagr2012=(value / value[year==2012])^(1 / (year - 2012)) * 100 - 100,
         cagr2016=(value / value[year==2016])^(1 / (year - 2016)) * 100 - 100)


cal_oes %>%
  filter(soc=="333012", wtype=="annual", atype=="state") %>%
  ggplot(aes(year, wage)) +
  geom_line() +
  geom_point()

dflong %>%
  ggplot(aes(year, ivalue, colour=name)) +
  geom_line() +
  geom_point() +
  scale_y_continuous(breaks=seq(-50, 50, 5)) +
  geom_hline(yintercept = 0)

dflong %>%
  ggplot(aes(year, cagr2016, colour=name)) +
  geom_line() +
  geom_point() +
  scale_y_continuous(breaks=seq(-50, 50, 0.5)) +
  geom_hline(yintercept = 0)

dflong %>%
  filter(year >= 2012) %>%
  ggplot(aes(year, cagr2012, colour=name)) +
  geom_line() +
  geom_point() +
  scale_y_continuous(breaks=seq(-50, 50, 0.5)) +
  geom_hline(yintercept = 0)


```





# Descriptive statistics from the ACS
```{r acs_factors}
# UPDATE THIS
# RAC1P 1
# Recoded detailed race code
# 1 .White alone
# 2 .Black or African American alone
# 3 .American Indian alone
# 4 .Alaska Native alone
# 5 .American Indian and Alaska Native tribes specified; or American
# .Indian or Alaska Native, not specified and no other races
# 6 .Asian alone
# 7 .Native Hawaiian and Other Pacific Islander alone
# 8 .Some Other Race alone
# 9 .Two or More Races
f_rac1p <- function(rac1p) {
  levs <- c(1, 2, 6, 3:5, 7:9)
  labs <- c("White alone", "Black alone", "Asian alone", "Am-Indian alone", "Alaska Native alone",
            "Am-Ind or AK native tribe", "Hawaiian or PacIsland alone", "Other race alone", "Two or more races")
  #factor(rac1p, levels=levs, labels=labs, ordered=TRUE)
  ordered(rac1p, levels=levs, labels=labs)
}

f_race <- function(rac1p) {
  levs <- c(1, 2, 6, 3:5, 7:9)
  x <- factor(rac1p, levels=levs, ordered = TRUE)
  levels(x) <- list("White alone"=1, 
                    "Black alone"=2, 
                    "Asian alone"=6, 
                    "Other race alone"=c(3:5, 7:8), 
                    "Two or more races"=9)
  return(x)
}
# table(f_race(c(1:9, 3, 5)))

f_race_hisp <- function(rac1p, hisp) {
  y <- ifelse(hisp!=1, 0, rac1p)
  levs <- c(1, 2, 6, 3:5, 7:9)
  y <- factor(y)
  levels(y) <- list("White alone, not Hispanic"=1, 
                    "Black alone, not Hispanic"=2,
                    "Hispanic, any race"=0,
                    "Asian alone, not Hispanic"=6,
                    "Other or multiple race, not Hispanic"=c(3:5, 7:9))
  return(y)
}

# (tmp <- f_race_hisp(rac1p=c(1, 1, 2, 2, 6), hisp=c(1, 2, 2, 1, 1)))
# table(tmp)

f_occgrp <- function(occp) {
  occgrp <- ifelse(occp=="6230", "carpenter", "other")
  occgrp <- ifelse(occp >= "6200" & occp <= "6940" & occgrp=="other", "construction", occgrp)
  occgrp <- ifelse(occp >= "3600" & occp <= "4650" & occgrp=="other", "service", occgrp)
  return(occgrp)
}

# f_occgrp(c("6230", "6231", "3610", "9000"))
```


```{r get_acs}
dbDisconnect(db) # in case it is connected - will throw an error if not, but that's ok
db <- dbConnect(SQLite(), dbname=acsdb)

vars <- "serialno, st, adjinc, pwgtp, agep, cow, mar, schl, sex, wagp, hisp, rac1p, occp"

sqlcmd <- paste0("SELECT ", vars, " FROM pus WHERE occp=3802") # 10 secs
system.time(df <- collect(tbl(db, sql(sqlcmd))))  # 3 secs
glimpse(df)
count(df, st)

df2 <- df %>%
  left_join(stcodes %>% 
              select(st=stfips, stabbr) %>% 
              mutate(st=as.integer(st)),
            by="st") %>%
  filter(wagp > 0, agep >= 18) %>%
  mutate(adjinc=adjinc / 1e6)

sts <- c("CA", "MA", "RI", "CT", "NY", "FL", "NV", "TX") %>% sort
storder <- c("other", sts)
df3 <- df2 %>%
  filter(cow==4) %>%
  group_by(stabbr) %>%
  mutate(n=n()) %>%
  ungroup %>%
  mutate(stgroup=ifelse(stabbr %in% sts, stabbr, "other"),
         stgroup=factor(stgroup, levels=storder),
         stabbr20plus=ifelse(n < 20, "LT20", stabbr),
         agep2=agep^2,
         exp=agep - 18,
         exp2=exp^2,
         lwage=log(wagp),
         sex=factor(sex, levels=1:2, labels=c("male", "female")),
         edattain=case_when(schl <= 15 ~ "notHSgrad",
                            schl %in% 16:19 ~ "HSgrad",
                            schl == 20 ~ "associate",
                            schl == 21 ~ "BA",
                            schl %in% 22:24 ~ "MA+",
                            TRUE ~ "other"),
         edattain=factor(edattain, levels=c("notHSgrad", "HSgrad", "associate", "BA", "MA+", "other")),
         marstat=factor(mar, levels=1:5, labels=c("married", "widowed", "divorced", "separated", "neverM")),
         marstat=factor(mar, levels=1:5, labels=c("marwid", "marwid", rep("notmarwid", 3))))
# count(df3, stgroup, stabbr)
count(df3, stgroup, edattain)
count(df3, stabbr)


mod <- lm(lwage ~ sex + stgroup + exp + exp2 + edattain + marstat, data=df3)
summary(mod)

mod1 <- lm(lwage ~ sex + stabbr + exp + exp2 + edattain + marstat, data=df3)
summary(mod1)

# https://cran.r-project.org/web/packages/fixest/fixest.pdf
# https://cran.r-project.org/web/packages/fixest/vignettes/fixest_walkthrough.html
# feols(log(Euros) ~ log(dist_km) | Origin + Destination + Product + Year, trade)
mod2 <- feols(lwage ~ sex + stabbr + exp + exp2 + edattain + marstat, data=df3)
summary(mod2)

mod3 <- feols(lwage ~ sex + exp + exp2 + edattain + marstat | stabbr20plus, data=df3)
summary(mod3)

fe <- fixef(mod3)
summary(fe)
fe
fe$stabbr20plus - mean(fe$stabbr20plus)
tibble(stabbr=names(fe$stabbr20plus), fe=fe$stabbr20plus) %>%
  mutate(cfe=fe - mean(fe),
         stname=get_stname(stabbr)) %>%
  ggplot(aes(y=reorder(stname, cfe), x=cfe * 100)) +
  geom_col(fill="blue") +
  theme_minimal() +
  labs(x="% above or below average state", y=NULL) +
  ggtitle("Correctional officer salaries relative to U.S. average correctional officer salaries",
          subtitle = "After controlling for experience, education, and demographic factors") +
  theme(axis.text.y = element_text(hjust = 0))

df <- tibble(stabbr=names(fe$stabbr20plus), fe=fe$stabbr20plus) %>%
  mutate(stabbr=ifelse(stabbr=="LT20", "Other", stabbr),
         cfe=fe - mean(fe),
         stname=get_stname(stabbr),
         stname=ifelse(stabbr=="Other", "Other", stname))

df %>%
  filter(stabbr != "Other") %>%
  mutate(grp=ifelse(stabbr=="CA", "CA", "other")) %>%
  ggplot(aes(y=reorder(stname, cfe), x=cfe * 100)) +
  geom_point(aes(colour=grp), size=2) +
  scale_colour_manual(values=c("red", "blue")) +
  theme_minimal() +
  labs(x="% above or below average state", y=NULL) +
  ggtitle("State correctional officer salaries relative to those in the average state",
          subtitle = "After controlling for experience, education, and demographic characteristics") +
  theme(axis.text.y = element_text(hjust = 0)) +
  geom_vline(xintercept = 0) +
  scale_x_continuous(breaks=seq(-100, 100, 10)) +
  theme(legend.position = "none")

  
 
# ggplot(aes(y=reorder(stname, change), x=change)) +
plot(fe)
plot(fe, n=51)
fe$stabbr
coef(mod3)
coeftable(mod3)
confint(mod3)
confint(mod3, se="cluster")  # same, because default is stabbr clustering
etable(mod3)
fitstat(mod3, type = "r2")
fitstat(mod3, type = "rmse")


mod4 <- feols(lwage ~ sex + exp + exp2 + edattain + marstat | stabbr, weights=df3$pwgtp, data=df3)
summary(mod4)

# MAR Character 1
# Marital status
# 1 .Married
# 2 .Widowed
# 3 .Divorced
# 4 .Separated
# 5 .Never married or under 15 years old


```



## Corrections officers
```{r}

```



# Notes about other studies

## Hedonic wage regressions
```{r eval=FALSE}
# Biggs Illinois
# The sample is limited to individuals who work for the state government or for the private sector; federal employees or employees of non-profits are excluded. The sample also is limited to individuals who work 35 or more hours per week and 50 or more weeks per year. The model used here controls for: Years of education; Undergraduate degree field (for those with a college education and above); Potential work experience (equal to age minus years of education minus 6) and experience-squared; Broad occupation (eight categories); place of residence (based upon the Census Bureau’s Public Use Microdata Areas, or PUMAs); Usual hours of work per week; Gender; Race; Hispanicity; Marital status; Immigrant status; Year; and State government employee. Our model has several key differences that can influence the results. First, we control for the number of hours worked per week. Since Illinois state government employees work about 1.5 fewer hours per week than private sector workers, excluding weekly hours worked makes state government employees appear relatively less well paid. Second, we include a variable for the employee’s undergraduate college major, acknowledging that certain majors lead to higher pay in the workforce than others….Third, we include data on the location within the state in which the employee lives.

# Biggs Illinois
# Wages - hedonic (human capital) - ACS controls vs above:
#   state gov penalty of about 2%, w/no firm size adjustment (add'l penalty of 7.2% if firm size adjust)


```


## Benefits

```{r}
# Biggs Illinois
# p.4 notes they received unpublished ECEC in Biggs and Richwine (2014),
#   for private workers and a small number of public benefits
#   HC, RHC, and pensions drawn from specific sources
# HC:
#   IL -- Pew 2014 -- about 42% higher than private (implies ~ 20.2% of salaries I think)
#   private -- ECEC -- about 14.2% of salaries (must be for region)
# RHC:
#   IL -- NC from AV, div salaries, gives 19.4% 
#   private --  (not in the ECEC)
#     based on Biggs Richwine, which adjusted CBO, data -- nationally 0.5% NC, IL 0.48%
# Pensions:
#   IL -- tot NC of 20.9% adjusted to 41.8% using DR of 5.0% vs. earnings assump of 7.25% (appears
#         to assume 32.7 year - perhaps full career?), less eec of 5.4%,
#         leaving er NC of 36.4%
#   private -- ECEC -- er contribs of 2.6 to DC + 1.4 to DB = 4.0 erc, vs 36.4% public
# Other FB (paid leave - vacation, holiday, personal time); emp prem for life, DB; legally req
#           benefits wf wno AA rZWAM nXewm YUm QX:
#   IL -- ECEC (for region)
#   private -- ECEC (for region)

# Biggs Richwin
# HC: 
  # Our principal source is data compiled by the National Conference of State
  # Legislatures (NCSL).37 These data show monthly employer and employee health
  # premiums for individual and family coverage in 2012. In certain cases, data
  # for 2012 were not available through the NCSL. In these cases we used NCSL
  # data from prior years, adjusted upward by the rate of growth of overall
  # health premiums through 2012. In several other cases, we obtained data
  # directly from state sources, such as budget documents.
#  NCS useful


```


